Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-30 paper.tex
some notes on performance

            Scientific workflows historically relied on I/O to transfer data
            between nodes. As scientific workflows are usually executed on
            HPC infrastructure, this meant writing to a filesystem such as 
            Lustre

            \subsection{Lustre}
            \subsection{Hadoop Distributed File System}
            \subsection{Amazon S3}

            scientific: shared FS such as Lustre, NFS, S3
            big data: mapreduce - HDFS, spark and dask shared or distributed
        \section{Scheduling}\label{sched}
            scientific: SLURM, PBS, SGE, HTCondor, Mesos (sometimes)
            big data: mapreduce - local or YARN, Spark - standalone, YARN or 
            Mesos, dask: any (easy to extend to any scheduler)
        \section{Performance enhancement strategies}\label{other}
            scientific: workflow checkpointing (looking at file hashes)
            big data: data locality, in memory computing, lazy evaluation
    \chapter{Portability}\label{portability}
        java or python backend (most) typically platform independent (some 
        exceptions)
        all tools used in the pipeline must be installed on system
        Taverna - exception as relies on web services
        \section{Containers}
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        \section{Provenance Tracking}
        big data frameworks: limited to lineage
        \section{containerization}
    \chapter{Discussion/Conclusion}

>>>>

            Unlike MapReduce and Spark-based workflows, Dask and scientific 
            workflows are not data-centric. They focus on the efficient 
            scheduling of tasks and rely on I/O and network transfers to
            communicate the data between nodes. With BigData workflows, such
            operations could incur significant performance penalties. 
            Filesystems. To mitigate this and remain efficient in the face of
            BigData, Dask implements lazy evaluation and in-memory processing.
            BigData platforms such as MapReduce rely heavily on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network. Spark
            uses data locality, lazy evaluation and in-memory computing to 
            ensure its optimal performance.
            For the efficient processing of neuroimaging BigData, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

            \subsection{Strategies}
                \subsubsection{in-memory computing}
                    Random Access Memory (RAM) is the fastest memory available
                    to workflow engines. BigData frameworks, such as Spark and 
                    Dask, leverage this feature by maintaining data in memory 
                    between tasks rather that saving them to local disk or 
                    shared memory which may incur significant overhead.

                    Spark and Dask can both read from filesystems available to
                    all nodes. When an action (Spark) or a compute (Dask) is 
                    called, the data is read from the storage and uploaded into
                    memory. Tasks use the data located in memory to compute the
                    results. Should data not fit in memory, the results may 
                    spill to local disk, in the instance where nodes contain
                    storage. Otherwise, the engines will spill data to shared 
                    filesystem. The later is a common case in HPC systems as 
                    storage and compute resources are typically separate, 
                    however, these shared filesystems are typically slower than 
                    local storage, and therefore, there will be additional 
                    performance penalties.

                    As data is not stored within a resilient distributed 
                    filesystem such as Lustre and HDFS, it risks being 
                    permanently lost if a node failure occurs. Spark and Dask
                    have both addressed this through the notion of 
                    \textit{lineage}. Lineage refers to the set of operations
                    that the data has undergone. By recording the lineage, lost
                    data will always be recoverable as the set of operations
                    to reproduce the data are known. Should a node be lost to 
                    the cluster, Spark and Dask reschedule the tasks required 
                    to rebuild the dataset to other nodes.

                    \note{mention caching}

                    While Spark and Dask both rely on the concept of shared 
                    memory, Spark avoids data transfers as often as possible by
                    scheduling tasks to nodes where the memory is located. Dask
                    , however, will schedule its tasks to the next available 
                    node.

                \subsubsection{Data locality}
                    Data locality, with respect to BigData frameworks, is a
                    scheduling
                    strategy in which location of the data is known by the 
                    scheduler and tasks are scheduled to the location where the
                    necessary data is located. It was one of the first main 
                    strategies implemented to make BigData processing more 
                    efficient. The reasoning behind data locality is that 
                    network transfers are expensive, particularly within 
                    commodity infrastracture, and therefore transferring large 
                    volumes of data across the network for every task would be 
                    rather expensive.

                    While MapReduce and Spark both partition data across the 
                    network similarly, they achieve data locality slightly 
                    differently. MapReduce can use the local filesystem in 
                    standalone mode, but requires use of a distributed 
                    filesystem executing on the compute nodes such as the 
                    Hadoop Distributed File System (HDFS). The reason as to why
                    parallel file systems such as Lustre do not work is that 
                    the storage nodes on such file systems are separate from 
                    the compute nodes. Separate storage and compute nodes 
                    eliminate any possiblity of data locality as network 
                    transfers will occur regardless of storage node. When the 
                    next MapReduce task is available, the master node will 
                    query the distributed file system to determine the 
                    available nodes closest. The master will first attempt to
                    schedule the task on a node containing the data, but if 
                    that node is not available, it will attempt to schedule
                    the task on the same rack. If all else fails, the task will
                    be scheduled on a node found on a different rack.

                    In contrast to MapReduce, Spark does not require data to 
                    be stored on a HDFS-like filesystem. It can read data from
                    any form of shared-filesystem, including Lustre. If data 
                    is extracted from a non HDFS-like filesystem, the initial 
                    loading of data will not use data locality. However, in all
                    future tasks in which the data is already loaded into an 
                    RDD, the scheduler will be able to assign task to the most
                    favorable partition dictated by delay scheduling.

                \subsubsection{lazy evaluation}

                    Lazy evaluation is a strategy used throughout computer 
                    science to improve performance of operations. Rather than
                    evaluating a expression immediately, it will wait until it 
                    is required in order to compute the result. This is 
                    achieved in Spark through the calling of an \textit{action}
                    on an RDD and in Dask through the calling of a 
                    \textit{compute} function on one of its data structures.
                    By using lazy evaluation, the engines are capable of 
                    constructing graphs that result in the most optimal
                    task order of execution.

            \subsection{BigData Filesystem optimizations for HPC}

                Writting BigData to Lustre is very costly compared to local 
                node storage. BigData filesystems like HDFS may not necessarily
                be the best solution for HPC as they will not only use up 
                storage on compute nodes, but also replicate the data across 
                the cluster taking up a significant amount of storage space.
                Moreover, while BigData frameworks typically execute on a 
                dedicated BigData cluster, they will be on a shared cluster in
                HPC.

                Alluxio, formerly known as tachyon, is an in-memory filesystem
                that stores lineage data rather than the data itself. It is
                assumed that recomputation of data will occur more rapidly than
                data transfer over the network. However,

                Triple-H leverages the heterogeneity of HPC clusters to 
                optimize the data's placement across the cluster for most 
                efficient use.

            \note{scientific: shared FS such as Lustre, NFS, S3
            big data: mapreduce - HDFS, spark and dask shared or distributed}
        \section{BigData Scheduling}\label{sched}

            Scientific workflows typically rely on HPC schedulers for 
            distributed scheduling on clusters. These schedulers include SLURM,
            PBS,TORQUE, SGE, HTCondor and Mesos. Additionally, they may provide
            their own form of scheduling for computations on a single node. 
            Nipype uses mutliproc... etc. BigData frameworks, such as Dask,
            also provide interfaces to these HPC schedulers. However, other
            frameworks such as MapReduce and Spark are limited to BigData 
            clusters such as Yet Another Resource Negotiator (YARN and MESOS).
            Spark additionally provides its own standalone scheduler. Similarly
            to scientific workflows, all three BigData frameworks provide their
            own schedulers for single-node scheduling.

            Spark's standalone scheduler is a simple scheduler that can be used
            in distributed workflows. It can execute in two modes: client and
            cluster. In client mode, the  \note{cannot run in cluster mode in pyspark}

            YARN needs hadoop cluster

            Mesos can run on HPC

            dask distributed
            \note{scientific: SLURM, PBS, SGE, HTCondor, Mesos (sometimes)
            big data: mapreduce - local or YARN, Spark - standalone, YARN or 
            Mesos, dask: any (easy to extend to any scheduler)}
        \section{Performance enhancement strategies}\label{other}
            scientific: workflow checkpointing (looking at file hashes)
            big data: data locality, in memory computing, lazy evaluation
    \chapter{Portability}\label{portability}
        java or python backend (most) typically platform independent (some 
        exceptions)
        all tools used in the pipeline must be installed on system
        Taverna - exception as relies on web services
        \section{Containers}

             Containeri
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        \section{Provenance Tracking}
        big data frameworks: limited to lineage
        \section{containerization}
    \chapter{Discussion/Conclusion}
